---
title: "Final Project"
author: "Abhishek Pathak"
date: November 14th 2019
output:
  
  word_document: default
---

# file retrival
```{r}
student <- read.csv("C:/Users/user/Documents/StudentsPerformance.csv", header = TRUE, stringsAsFactors = FALSE)
```

#examining the dataset
```{r}
str(student)
```

```{r}
head(student)
```

```{r}
summary(student)
```

```{r}
tail(student)
```

```{r}
class(student)
```

install.packages("dplyr")

library(dplyr)

install.packages("magrittr")

library(magrittr)

#Renaming the column names 
```{r}
colnames(student)
namesofcolumns<- c("Gender","Race","Parent_Education","Lunch","Test_Preparation","Math_score","Reading_score","Writing_score")
colnames(student)<- namesofcolumns
str(student)
```

#Removing the "Race" column from dataset 
```{r}
colnames(student)[2]<- "Race"
student<- subset( student, select = -Race)
```

#checking and remvoing N/A's from dataset
```{r}
sum(is.na(student$Gender))
sum(is.na(student$Parent_Education))
sum(is.na(student$Lunch))
sum(is.na(student$Test_Preparation))
sum(is.na(student$Math_score))
sum(is.na(student$Reading_score))
sum(is.na(student$Writing_score))
```

table(student$Gender)
table(student$Parent_Education)

head(student)

library(ggplot2)

#Histrogram for Math score and Gender
```{r}
a<- ggplot(student, aes(x = student$Math_score, fill=Gender)) + geom_histogram(binwidth = 4, color ="Red") + xlab("Math Scores") + ylab("Gender") + ggtitle("Math scores divided by Gender")
a
```

#Histrogram for Reading score and Gender
```{r}
b<-ggplot(student, aes(x = student$Reading_score, fill=Gender)) + geom_histogram(binwidth = 4, color ="Green") + xlab("Reading Scores") + ylab("Gender") + ggtitle("Reading scores divided by Gender")
b
```


#Histrogram for Reading score and Gender
```{r}
c<- ggplot(student, aes(x = student$Writing_score, fill=Gender)) + geom_histogram(binwidth = 4, color ="Black") + xlab("Writing Scores") + ylab("Gender") + ggtitle("Writing scores divided by Gender")
c
```

#Bar graph regarding number of students by gender
```{r}
d<- ggplot(student, mapping = aes(x = student$Gender)) + geom_bar(width = 0.3, fill = "red") + geom_text(stat = 'count', student, mapping =  aes(x = student$Gender, label =..count..)) + xlab("Gender") + ylab("Number of students") + ggtitle("Number of Students by Gender") + scale_fill_brewer(type = "qual", palette = 1, direction = 1, aesthetics = "fill") + ylim(0, 1200)
d
```

#Scatterplot - Math score vs Reading score
```{r}
e<- ggplot(data = student, aes(x= student$Math_score, y= student$Reading_score)) + geom_point() + xlab("Math Score") + ylab("Reading Score") + ggtitle("Relationship between Math Score and Reading Score of a Student")
e
```

#Scatterplot - Math score vs Writing score
```{r}
f<- ggplot(data = student, aes(x= student$Math_score, y= student$Writing_score)) + geom_point() + xlab("Math Score") + ylab("Writing Score") + ggtitle("Relationship between Math Score and Writing Score of a Student") 
f
```


#Bar grpah regarding Math score vs lunch
```{r}
h<- ggplot(student, aes(x= student$Lunch, y= student$Math_score)) + geom_bar(stat = "identity", fill = "orange") + xlab("Lunch") + ylab("Math Score") + ggtitle("How does Lunch affect Math score") + ylim(0, 100)
h
```

#Bar grpah regarding Reading score vs lunch
```{r}
i<- ggplot(student, aes(x= student$Lunch, y= student$Reading_score)) + geom_bar(stat = "identity", fill = "Yellow") + xlab("Lunch") + ylab("Reading Score") + ggtitle("How does Lunch affect Reading score") + ylim(0, 100)
i
```

#Bar grpah regarding Writing score vs lunch
```{r}
j<- ggplot(student, aes(x= student$Lunch, y= student$Writing_score)) + geom_bar(stat = "identity", fill = "Red") + xlab("Lunch") + ylab("Writing Score") + ggtitle("How does Lunch affect Writing score") + ylim(0, 100)
j
```

str(student)
sum(is.na(student))

###

#For the training and testing sets 70% of the total records will be training
#While the remaining 30% will be atributted towards testing
```{r}
student_train<- student[1:1400,]
student_test<- student[1400:2000,]
```

#Math score regression model
```{r}
model_math<- lm(student$Math_score ~ student$Gender + student$Parent_Education + student$Lunch + student$Test_Preparation + student$Reading_score + student$Writing_score, data = student_train)

summary(model_math)

predict_model_math<- predict(model_math, newdata = student_test, interval = "pred")
summary(predict_model_math)
```

#Reading score regression model
```{r}
model_reading<- lm(student$Reading_score ~ student$Gender + student$Parent_Education + student$Lunch + student$Test_Preparation + student$Math_score + student$Writing_score, data = student_train)

summary(model_reading)

predict_model_reading<- predict(model_reading, newdata = student_test, interval = "pred")
summary(predict_model_reading)
```

#Writing score regression model
```{r}
model_writing<- lm(student$Writing_score ~ student$Gender + student$Parent_Education + student$Lunch + student$Test_Preparation + student$Math_score + student$Reading_score, data = student_train)

summary(model_writing)

predict_model_writing<- predict(model_writing, newdata = student_test, interval = "pred")
summary(predict_model_writing)
```


